### 2020-09-09 paper/iros_abstract.tex
IEEE formatted paper

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.

% Bibliography
\usepackage{biblatex}
\addbibresource{references.bib}

% Math
\usepackage{physics}
\usepackage{siunitx}
\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

% Optimization and Algorithms
\usepackage{optidef}
\usepackage{algorithmicx}
\usepackage{algorithm,algpseudocode}

% Formatting
\usepackage{xcolor}
\usepackage{bm}  % for bold symbols 
\usepackage{booktabs}  % better tables
\usepackage{pifont}  % for x mark
\usepackage{graphicx}
\usepackage{hyperref}

% Plotting
\usepackage{pgfplots}
\pgfplotsset{compat=1.15,
	legend style={font=\footnotesize},
}
\usepackage{tikzscale}

% Custom commands
\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{S}^3}
\newcommand{\skewmat}[1]{[#1]^\times}

\newcommand{\rmap}{\varphi}
\newcommand{\invrmap}{\varphi^{-1}}

\newcommand{\dR}{\delta \mathcal{R}}
\newcommand{\rot}{ \mathcal{R} }
\newcommand{\dq}{\delta q}
\newcommand{\q}{\textbf{q}}
\newcommand{\eq}{_\text{eq}}
\newcommand{\traj}[2][N]{#2_{0:{#1}}}
\newcommand{\pass}{{\color{green} \checkmark}}
\newcommand{\fail}{{\color{red} \ding{55}}}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

\title{\LARGE \bf
Planning with Attitude
}

\author{Brian Jackson$^1$ and Zachary Manchester$^1$%
    \thanks{
        $^1$Robotics Institute, 
        Carnegie Mellon University, 
        5000 Forbes Ave, Pittsburgh, PA, USA
    }
}

\begin{document}
\maketitle

\begin{abstract}
Planning and controlling trajectories for floating-base robotic systems that experience
large attitude changes is challenging due to the nontrivial group structure of 3D
rotations. This paper introduces an accessible and unified approach for tracking control
and optimization-based planning on the space of rotations based on vector calculus and
linear algebra. The methodology is used to derive an extension of the Linear-Quadratic
Regulator (LQR) to systems with arbitrary attitudes, which we call Multiplicative LQR
(MLQR). We compare MLQR to a state-of-the-art geometric tracking controller designed
specifically for quadrotors and demonstrate superior tracking and robustness. An
iterative variant of MLQR is compared to existing methods on useveral constrained
trajectory optimization problems, again demonstrating improved runtime performance. %to
optimize trajectories for a variety of robotic systems, we provide benchmark comparisons
between several of the most common attitude representations used for motion planning and
control and find that the combination of unit quaternion state variables with Rodrigues
parameters to represent attitude errors provides an excellent combination of performance
and algorithmic simplicity.
\keywords{
    motion planning and control, quaternions, optimal control, linear quadratic regulator
}
\end{abstract}

\section{INTRODUCTION}
    Many useful robotic systems---including quadrotors, airplanes, satellites, autonomous 
    underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional 
    translations and rotations as part of their normal operation. 
    While representing translations is straightforward and intuitive, effectively 
    representing the nontrivial group structure of 3D rotations has been a topic of study 
    for many decades. Although we can intuitively deduce that rotations are three-dimensional, 
    a globally non-singular three-parameter representation of the space of rotations does 
    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing rotations, 
    we must either a) pick a three-parameter representation and deal with discontinuities, 
    or b) pick a higher-dimensional representation and deal with constraints between the 
    parameters. While simply representing attitude is nontrivial, generating and tracking 
    motion plans for floating-base systems is an even more challenging problem.

    Early work on control problems involving the rotation group dates back to the 1970s, 
    with extensions of linear control theory to spheres \cite{Brockett1973} and $SO(3)$ 
    \cite{Baillieul1978}. Effective attitude tracking controllers have been developed for 
    satellites \cite{wie1985quaternion}, quadrotors \cite{Fresk2013,Liu2015,lee2010geometric,
    Johnson2005,watterson2020control,mellinger2011minimum}, and a 3D inverted pendulum 
    \cite{Chaturvedi2009}  using various methods for calculating three-parameter attitude 
    errors.

    More recently, these ideas have been extended to trajectory generation \cite{Zefran1998}, 
    sample-based motion planning \cite{Zefran1999,Kuffner2004}, and optimal control. 
    Approaches to optimal control on attitude problems include analytical methods applied to
    satellites \cite{Spindler1998}, discrete mechanics \cite{Kobilarov2011,Kobilarov2014,
    Lee2008}, a combination of sampling-based planning and constrained trajectory 
    optimization for satellite formations \cite{Garcia2005, Aoude2008}, projection operators 
    \cite{Saccon2013}, or more general theory for optimization on manifolds 
    \cites{watterson2018trajectory}. 
    Nearly all of these methods rely heavily on principles from differential geometry and 
    Lie group theory; however, despite these works, many recent papers in the robotics 
    community continue to apply traditional methods for motion planning and control with no 
    regard for the group structure of rigid body motion \cite{Alothman2016,deCrousaz2015,
    Williams2017,Geisert2016}.

    In this paper, we make a departure from previous approaches to geometric planning and 
    control that rely heavily on ideas and notation from differential geometry, 
    and instead use only basic mathematical tools from linear algebra and calculus that 
    should be familiar to most roboticists. 
    Similar to \cite{Mandic2011,Xu2016}, in Sec. \ref{sec:Quaternion_Calculus} we introduce 
    a quaternion differential calculus, but take a significantly simpler and more general 
    approach that unifies both planning and control, enabling straight-forward adaptation of 
    existing algorithms to systems with quaternion states. 
    To make this concrete, in Sec. \ref{sec:MLQR} we apply our approach to derive an 
    extension to the linear-quadratic regulator (LQR) that we call multiplicative LQR (MLQR), 
    which is the control dual to the multiplicative extended Kalman filter (MEKF) from the 
    state estimation literature \cite{markley2014fundamentals}. 
    In Sec. \ref{sec:experiments} we provide several simulation results demonstrating the 
    application of our method to tracking control and constrained trajectory optimization.
    In summary, our contributions include:

    \begin{itemize}
        \item A unified approach to quaternion differential calculus entirely based on standard vector calculus and linear algebraic operations
        \item Derivation of multiplicative LQR, an adaptation of LQR to the control of systems with quaternion states
        \item A set of benchmark motion-planning and trajectory-tracking problems in which we compare our approach to existing methods commonly used in the robotics community to account for 3D rotations
    \end{itemize}

\section{Background}

    We begin by defining some useful conventions and notation. 
    Attitude is defined as the rotation from the robot's body frame to a global inertial 
        frame. 
    We also define gradients to be row vectors, that is, for 
        $f(x) : \R^n \to \R$, $\pdv{f}{x} \in \R^{1\times n}$.

    \subsection{Unit Quaternions} \label{sec:quaternions}
        We leverage the fact that quaternions are linear operators and that the space of 
        quaternions $\mathbb{H}$ is isomorphic to $\R^4$ to explicitly 
        represent---following the Hamilton convention---a quaternion $\q \in \mathbb{H}$ as 
        a standard vector $q \in \R^4 := [q_s \;\; q_v^T]^T$ where $q_s \in \R$ and 
        $q_v \in \R^3$ are the scalar and vector part of the quaternion, respectively.

        Quaternion multiplication is defined as
        \begin{equation} \label{eq:quat_mult}
            \q_2 \otimes \q_1 = L(q_2) q_1 = R(q_1) q_2
        \end{equation}
        where $L(q)$ and $R(q)$ are orthonormal matrices defined as
        \begin{align}
            L(q) &:= \begin{bmatrix} 
                q_s \;\; & -q_v^T \\ 
                q_v \;\; & q_s I + \skewmat{q_v} 
            \end{bmatrix} 
            \label{eq:Lmult} \\
            %= \begin{bmatrix} q & G(q) \end{bmatrix} \label{eq:Lmult} \\
            R(q) &:=\begin{bmatrix} 
                q_s \;\; & -q_v^T \\ 
                q_v \;\; & q_s I - \skewmat{q_v} 
            \end{bmatrix} \label{eq:Rmult},
        \end{align}
        and $\skewmat{x}$ is the skew-symmetric matrix operator
        \begin{equation}
            \skewmat{x} := \begin{bmatrix} 
                0 & -x_3 & x_2 \\ 
                x_3 & 0 & -x_1\\ 
                -x_2 & x_1 & 0 
            \end{bmatrix}.
        \end{equation}

        The inverse of a unit quaternion $q^{-1}$, giving the opposite rotation, is equal 
        to its conjugate $q^*$, which is simply the same quaternion with a negated vector 
        part:
        \begin{equation} \label{eq:T}
            \q^* = T q := \begin{bmatrix} 
                1 & \\ 
                & -I_3 
            \end{bmatrix} q
        \end{equation}
        The following identities, which are easily derived from 
        \eqref{eq:Lmult}--\eqref{eq:T}, are extremely useful:
        \begin{align}
            &L(Tq) = L(q)^T = L(q)^{-1} \\
            &R(Tq) = R(q)^T = R(q)^{-1} .
        \end{align}

        We will sometimes find it helpful to create a quaternion with zero scalar part from 
        a vector $r \in \R^3$. We denote this operation as,
        \begin{equation}
            \hat{r} = H r \equiv \begin{bmatrix} 0 \\ I_3 \end{bmatrix} r.
        \end{equation}
        Unit quaternions rotate a vector through the operation 
        $\hat{r}' = \q \otimes \hat{r} \otimes \q^*$. 
        This can be equivalently expressed using matrix multiplication as
        \begin{align} 
            r' &= H^T L(q) R(q)^T H r = A(q)r , \label{eq:quaternion_rotation}
        \end{align}
        where $A(q)$ is the rotation matrix in terms of the elements of the quaternion 
        \cite{markley2014fundamentals}.

    \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
        In the current work we will restrict our focus to rigid bodies moving freely in 3D 
        space. That is, we consider systems with dynamics of the following form:
        \begin{equation} \label{eq:rigid_body_dynamics}
            x = \begin{bmatrix} r \\ R \\ v \\ \omega \end{bmatrix}, \quad 
            \dot{x} = \begin{bmatrix} 
                v \\ 
                \texttt{kinematics}(R,\omega) \\ 
                \frac{1}{m} F_G(x,u) \\ 
                J^{-1}(\tau_L(x,u) - \omega \times J \omega) 
            \end{bmatrix}
        \end{equation}
        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the position, 
        $R \in SO(3)$ is the attitude, $v \in \R^3$ is the linear velocity, and 
        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, 
        $J \in \R^{3\times3}$ is the inertia matrix, $F_G(x,u) \in \R^3$ are the forces in the 
        global frame, $\tau_L(x,u)$ are the moments in the local (body) frame, and 
        $\texttt{kinematics}$ are the kinematics for the chosen attitude representation. The 
        kinematics for unit quaternions are
        \begin{equation}
            \dot{q} = \half \q \otimes \hat{\omega} = \half L(q) H \omega.
        \end{equation}

\section{Quaternion Differential Calculus} \label{sec:Quaternion_Calculus}
    We now present a simple but powerful method for taking derivatives of functions 
    involving quaternions based on the notation and linear algebraic operations outlined 
    in Sec. \ref{sec:quaternions}.

    \begin{figure}
        \centering
        \includegraphics[height=5cm]{figures/tangent_plane.tikz}
        \caption{
            When linearizing about a point $q$ on an sphere $\mathbb{S}^{n-1}$ in 
            n-dimensional space, the tangent space $T$ is a plane living in $\R^{n-1}$, 
            illustrated here with $n=3$. Therefore, when linearizing about a unit 
            quaternion $q \in \Q$, the space of differential rotations lives in $\R^3$.
        }
        \label{fig:tangent_plane}
    \end{figure}

        Derivatives consider the effect an infinitesimal perturbation to the input has on
        an infinitesimal perturbation to the output. For vector spaces, the composition
        of the perturbation with the nominal value is simple addition and the
        infinitesimal perturbation lives in the same space as the original vector. For
        unit quaternions, however, neither of these are true; instead, they compose
        according to \eqref{eq:quat_mult}, and infinitesimal unit quaternions are (to
        first order) confined to a 3-dimensional plane tangent to $\Q$ (see Fig.
        \ref{fig:tangent_plane}).

        %, and are related through a non-linear mapping $\varphi : \R^3 \to \Q$ )
        The fact that differential unit quaternions are three-dimensional should make
        intuitive sense: Rotations are inherently three-dimensional and differential
        rotations should live in the same space as angular velocity, i.e. $\R^3$.

        skinny dog runs after skinny cat

        There are many possible three-parameter representations for small rotations in
        the literature. Many authors use the exponential map \cite{Baillieul1978,
        Zefran1998, Lee2008, Saccon2013, Sola2017, Fan2016, watterson2018trajectory},
        while others have used the Cayley map (also known as Rodrigues parameters)
        \cite{Kobilarov2011, Kobilarov2014}, Modified Rodrigues Parameters (MRPs)
        \cite{Terzakis2018}, or the vector part of the quaternion \cite{Fresk2013}.
        We choose Rodrigues parameters \cite{markley2014fundamentals} because they are
        computationally efficient and do not inherit the sign ambiguity associated with
        unit quaternions. The mapping between a vector of Rodrigues parameters $\phi \in
        \R^3$ and a unit quaternion $q$ is known as the Cayley map: \begin{equation}
        \label{eq:cayley}
            q = \varphi(\phi) = \frac{1}{\sqrt{1 + \norm{\phi}^2}} \begin{bmatrix} 1 \\ \phi \end{bmatrix}.
        \end{equation}
        We will also make use of the inverse Cayley map:
        \begin{equation}
            \phi = \varphi^{-1}(q) = \frac{q_v}{q_s}.
        \end{equation}

    \subsection{Jacobian of Vector-Valued Functions}
        When taking derivatives with respect to quaternions, we must take into account
        both the composition rule and the nonlinear mapping between the space of unit
        quaternions and our chosen three-parameter error representation.

        Let $\phi \in \R^3$ be a differential rotation applied to a function with
        quaternion inputs $y = h(q): \Q \to \R^p$, such that
        \begin{equation} \label{eq:vector_function}
            y + \delta y = h(L(q) \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
        \end{equation}
        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
        differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at
        $\phi = 0$:
        \begin{equation} \label{eq:quat_gradient}
            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) 
                        = \pdv{h}{q} \begin{bmatrix} 
                            -q_v^T \\ 
                            sI_3 + \skewmat{q_v}
                        \end{bmatrix}
        \end{equation}
        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which
        essentially becomes a ``conversion factor'' allowing us to apply results from
        standard vector calculus to the space of unit quaternions. This form is
        particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be
        obtained using finite difference or automatic differentiation.
        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the
        same (up to a constant scaling factor) for any choice of three-parameter attitude
        representation.

    \subsection{Hessian of Scalar-Valued Functions}
	    If the output of $h$ is a scalar ($p = 1$), then we can find its Hessian by
	    taking the Jacobian of \eqref{eq:quat_gradient} with respect to $\phi$ using the
        product rule, again evaluated at $\phi = 0$:

	    \begin{equation} \label{eq:quat_hessian}
            \nabla^2 h(q) = G(q)^T \pdv[2]{h}{q} G(q) + I_3 \pdv{h}{q}q,
	    \end{equation}
	    where the second term comes from the second derivative of $\varphi(\phi)$.
	    Similar to $G(q)$, this ends up being the same (up to a scaling factor) for any
        choice of three-parameter attitude representation.

    \subsection{Jacobian of Quaternion-Valued Functions}
        We now consider the case of a function that maps unit quaternions to unit
        quaternions, $q' = f(q) : \Q \to \Q$. Here we must also consider the non-trivial
        effect of a differential value applied to the output, i.e.:
        \begin{equation} \label{eq:dqoutput}
            L(q') \varphi(\phi') = f(L(q)\varphi(\phi)) .
        \end{equation}
        Solving \eqref{eq:dqoutput} for $\phi'$ we find,
        \begin{equation} \label{eq:phiprime}
            \phi' = \varphi^{-1} \left( L(q')^T f(L(q)\varphi(\phi)) \right) \approx \nabla f(q) \, \phi.
        \end{equation}
        Finally, the desired Jacobian is obtained by taking the derivative of
        \eqref{eq:phiprime} with respect to $\phi$:
        \begin{equation} \label{eq:quat_jacobian}
            \nabla f(q) = H^T L(q')^T \pdv{f}{q} L(q) H = G(q')^T \pdv{f}{q} G(q).
        \end{equation}
        The leading $G(q')^T$ comes from the fact that as $\phi' \to 0$, $L(q') f(q) \to
        I_q$, where $I_q$ is the quaternion identity. Differentiating through the inverse
        map, evaluated at the quaternion identity, we find that $\pdv*{\varphi^{-1}}{q}
        \to H^T$ for any three-parameter attitude representation.

\section{Multiplicative LQR} \label{sec:MLQR}
    Leveraging the methods from the previous section, we derive multiplicative LQR
    (MLQR), a variant of LQR that correctly accounts for the group structure of
    rotations. For concreteness, we consider a system with rigid body dynamics, as
    presented in Sec. \ref{sec:rigidbody_dynamics}, and design a controller to stabilize
    the system about a dynamically feasible discretized reference trajectory
    $\bar{x}_{0:N}, \bar{u}_{0:N}$ with $N$ time steps.
    We begin by linearizing the dynamics about the reference trajectory using \eqref{eq:quat_jacobian}. Our linearized error dynamics become
    \begin{equation} \label{eq:linearized_dynamics}
        \delta x_{k+1} = A_k \delta x_k + B_k \delta u_k 
    \end{equation}
    where \begin{equation}
        \begin{aligned}
            A_k = E(\bar{x}_{k+1})^T \pdv{f}{x}|_{\bar{x}_k,\bar{u}_k} E(\bar{x}_k), \\
            B_k = E(\bar{x}_{k+1})^T \pdv{f}{u}|_{\bar{x}_k,\bar{u}_k},
        \end{aligned}
    \end{equation}
    and $\delta x_k \in \R^{12}$ and $E(x_k) \in \R^{12 \times 13}$ are the state error and state error Jacobian:
    \begin{equation} \label{eq:state_error}
        \setlength\arraycolsep{1pt}
        \delta x_k = \begin{bmatrix} 
            r_k - \bar{r}_k \\ \varphi^{-1}(\bar{\q}_k^{-1} \otimes \q_k) \\ v_k - \bar{v}_k \\ \omega_k - \bar{\omega}_k 
        \end{bmatrix}\!, \;
        E(x) = \begin{bmatrix}
            I_3 & & & \\
            & G(q) & & \\
            & & I_3 & \\
            & & & I_3 \\
        \end{bmatrix}\!.
    \end{equation}

    With our linearized system, we can apply the traditional LQR Riccati recursion,
    \begin{multline}
         P_k = W_k +  A_k^T P_{k+1} A_k \\
         - A_k^T P_{k+1}^T B_k (R_k + B_k^T P_{k-1} B_k)^{-1} B_k^T P_{k+1} A_k ,
    \end{multline}
    which minimizes the quadratic cost function
    \begin{equation}
        \half \sum_{k=0}^N \delta x_k^T Q_k \delta x_k + \delta u_k^T R_k \delta u_k
    \end{equation}
    of the state \emph{error}, where $Q_k \in \R^{12 \times 12}$ is the weight matrix on
    the state \emph{error}, and $R_k \in \R^m$ is the weight matrix on the controls. From
    this, we get our non-linear feedback policy
    \begin{equation} \label{eq:mlqr_control}
        u = K_k \delta x_k + \bar{u}_k.
    \end{equation}
    where $\delta x_k$ is computed online using the non-linear error function
    \eqref{eq:state_error} and the linear feedback gain $K_k$ is calculated using the
    standard LQR formula:
    \begin{equation} \label{eq:LQR_gain}
         K_k := -(R_k + B_k^T P_{k+1} B_k)^{-1} B_k^T P_{k+1} A_k.
    \end{equation}

    The multiplicative Linear Quadratic Regulator is summarized as follows:
    \begin{enumerate}
        \item Pick $Q_k \in \R^{12 \times 12}$ to weight the state \textit{error}, and $R_k \in \R^{m \times m}$
        \item Linearize the dynamics using the state error Jacobian \eqref{eq:state_error}
        \item Compute $K_k$ using the standard LQR Riccati recursion
        \item Online, compute the control using \eqref{eq:state_error} and \eqref{eq:mlqr_control}.
    \end{enumerate}

    \subsection{Constrained Iterative MLQR}
        As demonstrated above, it is very straight-forward to adapt LQR to use
        quaternions. We similarly use this technique to adapt nonlinear trajectory
        optimization algorithms. Here we present the modifications to the ALTRO solver
        \cite{howell2019altro}, which uses iterative LQR (iLQR) combined with an
        augmented Lagrangian approach to handle constraints.

        For each iteration of iLQR within ALTRO, we can calculate a quadratic
        approximation of the cost function using \eqref{eq:quat_gradient} and
        \eqref{eq:quat_hessian}. Identical to \eqref{eq:linearized_dynamics}, we use
        \eqref{eq:quat_jacobian} to linearize the dynamics and use MLQR to solve for the
        feedback and feedforward gains.

        The only other modification to the algorithm is during the iLQR ``forward pass''
        that simulates the system with the closed-loop MLQR controller, where we use
        \eqref{eq:state_error} to calculate the nonlinear feedback policy during the
        forward roll-outs.
        For robustness, we may also add an additional criterion to the line search that
        checks for singularities in the three-parameter error state which, for the Cayley
        map, occur at \ang{180}, at which point the step size should be reduced.

    \subsection{Quaternion Cost Functions} \label{sec:cost_functions}
        In addition to the straight-forward modifications to the LQR algorithm itself, we
        need to carefully consider the types of cost functions we minimize. We frequently
        minimize costs that penalize distance from a goal state, e.g. $\half (x-x_g)^T Q
        (x-x_g)$; however, na\"ive substraction of unit quaternions is ill-defined. We
        propose two different cost functions that accomplish similar behavior.
        For sake of clarity and space, we only consider the costs on the quaternion
        variables: costs on the other states and the control variables remain unaffected.

        \subsubsection{Error Quadratic}
            Rather than simple subtraction, we can use a quadratic function on the
            three-parameter error state \eqref{eq:state_error}:
            \begin{equation} \label{eq:error_quadratic}
                J_\text{err} = \half \phi^T Q \phi 
                = \half \left(\varphi^{-1}(\delta q)\right)^T Q 
                        \left(\varphi^{-1}(\delta q)\right).
            \end{equation}
            where $\delta q = L(q_g)^T q$, and $\phi = \varphi{\delta q}$. The gradient
            and Hessian of \eqref{eq:error_quadratic} are
            \begin{equation}
                \nabla J_\text{err }= \phi^T Q  D(\delta q)  G(\delta q)
            \end{equation}
            \begin{multline}
                \nabla^2 J_\text{err} = 
                    G(\delta q)^T \! \left(
                    D(\delta q)^T Q D(\delta q) + \nabla D \right) G(\delta q)  \\
                    + I_3 (\phi^T Q  D(\delta q)) \delta q %\\
            \end{multline}
            where, for the Cayley map,
            \begin{equation}
                D(q) = \pdv{\varphi^{-1}}{q} = -\frac{1}{q_s^2}\begin{bmatrix}
                    q_v \;\; & -\frac{1}{q_s} I_3
                \end{bmatrix}
            \end{equation}
            \begin{equation}
                \nabla D = \pdv{q}(D(q)^T Q \phi) 
                = -\frac{1}{q_s^2} \begin{bmatrix} 
                    -2 \frac{q_v}{q_s}^T Q \phi & \phi^T Q \\
                                   Q \phi & 0 \\
                \end{bmatrix}.
            \end{equation}

        \subsubsection{Geodesic Distance}
            Alternatively, we can use the geodesic distance between two quaternions
            \cite{Kuffner2004},
            \begin{equation} \label{eq:quat_geodesic}
                J_\text{geo} = (1-\abs{q_g^T q}) ,
            \end{equation}
            whose gradient and Hessian are,
            \begin{align}
                \nabla J_\text{geo} &= \pm q_g^T G(q) \\
                \nabla^2 J_\text{geo} &= \pm I_3 q_g^T q ,
            \end{align}
            where the sign of the Hessian corresponds to the sign of $q_d^T q$.

\section{Experiments} \label{sec:experiments}
    Code for all experiments is available on 
    \href{https://github.com/RoboticExplorationLab/PlanningWithAttitude}
    {GitHub\footnote{\url{https://github.com/RoboticExplorationLab/PlanningWithAttitude}}}.

    \subsection{Trajectory Optimization}
        In this section we present several trajectory optimization problems solved using
        our modified version of ALTRO. The two quadrotor examples use 101 knot points for
        5 second trajectories, and the airplane example uses 51 knot points for a 1.25
        second trajectory. All results were run on a desktop computer with an Intel
        i9-9900 3.10 GHz processor with 32GB of memory. Timing results for all three
        experiments are given in Fig. \ref{fig:timing_chart}.

        In all examples, MLQR is implemented with a unit quaternion attitude state and
        Rodrigues parameters as the error state. The first example compares the two cost
        functions given in Sec. \ref{sec:cost_functions}, while the last two use the
        geodesic quaternion cost and compare iterative MLQR (iMLQR) from Sec.
        \ref{sec:MLQR} to iLQR using roll-pitch-yaw Euler angles following the convention
        from \cite{michael2010grasp}, and quaternions (treating them na\"ively as
        vectors). All examples re-normalize the quaternion in the dynamics function and
        use a third-order explicit Runge-Kutta integrator.

        \begin{figure}
            \centering
            \includegraphics[height=4cm, width=8cm]{figures/timing_chart.tikz}
            \caption{Timing results for the airplane barrell roll, quadrotor
                flip, and quadrotor zig-zag examples using iLQR with
                roll-pitch-yaw Euler angles (RPY), quaternions, and iMLQR. The
                timing result for the Quadrotor flip with Euler angles is omitted
                since it failed to converge.
            }
            \label{fig:timing_chart}
        \end{figure}

        \subsubsection{Quadrotor Zig-Zag}
            \begin{figure}[t]
                \centering
                \includegraphics[height=5.2cm]{figures/zig_zag_tracking.png}
                \caption{Zig-zag trajectory. The black quadrotors indicate
                    the reference trajectory generated by iMLQR. The yellow
                    quadrotors track the reference trajectory using time-varying
                    MLQR, while the red uses the HFCA controller from
                    \cite{watterson2020control}.
                }
                \label{fig:zig-zag}
            \end{figure}    
            \begin{figure}[t]
                \centering
                \includegraphics[height=5.2cm,trim={0 3cm 0 0},clip]{figures/quad_flip.png}
                \caption{Snapshots of the quadrotor flip trajectory. The
                    red-colored quadrotors represent the state near t=0 s and the
                    green-colored quadrotors represent the state near t=5.0 s
                }
                \label{fig:quad_flip}
            \end{figure}

        We set up a fairly simple trajectory optimization problem to achieve a
        ``zig-zag'' pattern, as shown in Fig. \ref{fig:zig-zag} using cost funtions of
        the form
        \begin{multline}
                  \half \sum_{k \in \mathcal{N}} \ell(x_k, x_\text{nom}, Q_\text{nom}) 
                + \half \sum_{k \in \mathcal{W}} \ell(x_k, \bar{x}_k, Q_w) \\
                + \half \sum_{k=1} u_k^T R u_k
        \end{multline}
        with $\mathcal{W} = \{33,66,101\}$, $\mathcal{N} = \{1:101\} \setminus
        \mathcal{W}$, and the waypoint positions were $(10,0,1)$,$(-10,0,1)$ and
        $(0,10,0)$. We compared the two cost functions from Sec.
        \ref{sec:cost_functions}, using a standard quadratic penalty on the
        non-quaternion states.
        \begin{align}
            \ell_\text{err}(x, \bar{x}, Q) &= \delta x^T Q \delta x \\ 
            \ell_\text{geo}(x, \bar{x}, Q) &= (x-\bar{x})^T \bar{Q} (x-\bar{x}) + w (1
                \pm \bar{q}^T q)
        \end{align}
        where $\bar{Q} = \text{diag}(Q_r, \vec{0}_4, Q_v, Q_\omega)$, where
        $Q_r,Q_v,Q_\omega$ are the weights of $Q$ for position, and linear and angular
        velocity.
        We used the following weighting matrices: 
        $Q_\text{nom} = \text{diag}(10^{-5} I_6,10^{-3} I_6)$, 
        $Q_w = \text{diag}(10^3 I_3,I_9)$, 
        $Q_N = \text{diag}(10 I_3, 100 I_3, 10 I_6)$.
        For the geodesic weights $w$, we used 0, 1, and 10 for the nominal, waypoint, and
        terminal cost functions, respectively.

        The geodesic cost function performed significantly better than the error
        quadratic, converging in 10 ms and 13 iterations, compared to 53 ms in 29
        iterations. In addition to better convergence behavior, the geodesic cost is
        about half as expensive to compute, which is unsurprising when comparing the
        gradients and Hessians in Sec. \ref{sec:cost_functions}.

        \subsubsection{Quadrotor Flip}
	    Similar to the previous example, we use a set of waypoints to encourage the
	    quadrotor to do a flip. The cost weights for this example were
	    \begin{align*}
	        Q_\text{nom} &= \text{diag}(10^{-2},10^{-2},5 \!\times\! 10^{-2}, 10^{-3} I_3,
    	        10^{-3} I_3, 10^{-2} I_3) \\
	        Q_w &= \text{diag}(10^3, 10, 10^3, 5000 I_3, I_3, 10 I_3) \\
	        Q_N &= \text{diag}(10 I_3, 100 I_3, 10 I_6)  
	    \end{align*}
	    Four intermediary knotpoints were used to encourage the quadrotor to be at angles
	    of \ang{90}, \ang{180}, \ang{270}, and \ang{360}. The quadrotor was constrained
	    to stay above the floor and move to a goal state 2 meters away in the $+y$
	    direction. The solver is initialized with a dynamically infeasible trajectory
	    that linearly interpolates between the initial and final states, rotating the
	    quad around the x-axis a full \ang{360}.

	    Figure \ref{fig:quad_flip} shows snapshots of the trajectory as generated using
	    iMLQR within the ALTRO solver to handle constraints. The trajectory generated by
	    iLQR with quaternions and quadratic costs was successful but had a slightly
        different trajectory. iLQR with Euler angles failed to solve the problem.

        \subsection{Airplane Barrel Roll}
        An airplane model with aerodynamic coefficients fit from real wind-tunnel data is
        tasked to do a barrel roll by setting a high terminal cost for being upside-down,
        see Fig. \ref{fig:barrellroll}. The solver is initialized with level flight trim
        conditions. For both the quadrotor flip and the airplane barrel roll, iMLQR
        converged faster than the pure quaternion version. Despite the extra matrix
        multiplications and extra Hessian term, it also was faster per iteration than its
        iLQR counterpart. For these highly aerobatic maneuvers, we achieve, as expected,
        better performance by correctly leveraging the structure of the rotation group
        during the optimization routine.

        \begin{figure}[h]
            \centering
            \includegraphics[height=3.5cm]{figures/barrellroll.png}
            \caption{Barrel roll trajectory computed by iterative MLQR using a terminal cost to encourage an upside-down attitude.}
            \label{fig:barrellroll}
        \end{figure}

	    \begin{figure}[h]
	        \centering
	        \includegraphics[width=\columnwidth,height=4.5cm]{figures/tracking_err.tikz}
	        \caption{Tracking error over time for the zig-zag trajectory using MLQR and the HFCA geometric controller. Band-limited white noise wrenches of variance $w$ are added to the continuous dynamics. }
	        \label{fig:tracking_err}
	    \end{figure}
        \vspace{-20pt}

	\subsection{Tracking Control}
	    This section provides examples using MLQR for tracking and stabilizing control
	    for a quadrotor. We compare our MLQR controller to HFCA, a state-of-the-art
	    geometric tracking controller \cite{watterson2020control} based on the Hopf
	    Fibration.

	    \subsubsection{MLQR Stabilization}
	    To test robustness and the domain of attraction of the MLQR controller, a
	    quadrotor is linearized about hover at the origin and initialized from states
	    uniformly sampled in $[-1,1]$ for position, $[-5,5]$ for linear and angular
	    velocities, and uniformly over the entire space of rotations. The success rates
	    for 2000 trials are given in Table \ref{tab:mlqr_success}, where success is
	    defined as the terminal error (calculated using \eqref{eq:state_error}) being
	    within a unit ball of radius 0.1 after 10.0 seconds. To further analyze the
	    basins of attraction, we uniformly sampled over only roll (which goes singular at
	    \ang{90} for our Euler angle representation), and yaw. The results are shown in
	    Fig. \ref{fig:mlqr_basin}, where the plotted contours are the convex hull of the
	    successfully stabilized initial conditions.

	    MLQR slightly exceeds the robustness of the HFCA geometric controller, while also
	    being more generally applicable, easier to tune, and easier to implement. It is
	    also noteworthy that na\"ively using LQR with quaternions (incorrectly treating
	    them as vectors) results in poorer performance than Euler angles, clearly
	    demonstrating the value of the methodology presented in the current paper.

	    \subsubsection{MLQR Tracking}
	    Time-varying MLQR is used to track the zig-zag trajectory generated in the
	    previous section. %The controller calculated the gains at $0.01 s$ intervals, and
	    used zero-order-hold between intervals. To test tracking performance, zero-mean
	    Gaussian noise was added as force and moments applied to the quadrotor.

	    The trajectories and tracking error for MLQR and HFCA are shown in Figs.
	    \ref{fig:zig-zag}, and \ref{fig:tracking_err}. While the MLQR controller
	    dramatically out-performs HFCA, we found no significant difference in performance
	    between MLQR and LQR using either roll-pitch-yaw Euler angles or a quaternion.
	    Since the tracking controller only deals with relatively small state and attitude
	    errors, this is unsurprising. Additionally, we found the MLQR controller ran
        about 20x faster than the HFCA controller.

	    \begin{figure*}[ht]
	        \begin{minipage}{0.60\linewidth}
                \centering
                \includegraphics[height=6.6cm,width=\textwidth]{figures/mlqr_basin.tikz}
                \caption{Basins of attraction of 4 stabilizing controllers for yaw and roll, calculated from 2000 sample points (shown in black). All points within each contour were successfully stabilized by the controller.}
                \label{fig:mlqr_basin}
            \end{minipage}
	        \hspace{0.05\linewidth}
            \begin{minipage}{0.34\linewidth}
                \centering
                \begin{tabular}{cc}
                    \toprule
                    \textbf{Method} & \textbf{Success} \\
                    \midrule
                    Quat & 61\% \\
                    RPY  & 80\% \\
                    HFCA & 97\% \\
                    MLQR & 99\% \\
                    \bottomrule
                \end{tabular}
                \caption{Success rate for Monte-Carlo analysis of quadrotor stabilization for 2000 initial conditions uniformly over position, orientation, and linear and angular velocities.}
                \label{tab:mlqr_success}
            \end{minipage}
        \end{figure*}

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for planning and control on rigid-body
    systems with arbitrary attitude using standard linear algebra and vector calculus. By
    applying these methods to LQR to correctly account for the group structure of
    rotations, we have matched or exceeded the performance of a state-of-the-art
    geometric controller designed specifically for quadrotors, while being more general,
    requiring less system-dependent tuning, having less computational overhead, and being
    easier to implement.

    We have also demonstrated a straight-forward way to adapt nonlinear trajectory
    optimization techniques to work with quaternion-valued states. For these problems, we
    found the geodesic distance between quaternions \eqref{eq:quat_geodesic} to be
    computationally efficient and provide excellent convergence in practice. We recommend
    the use of unit quaternions within planning and control algorithms for their
    simplicity, computational efficiency, and lack of singularities. We further recommend
    the use of Rodrigues parameters, or the Cayley map, as the quaternion error state.

    Future areas of research include the application of these methodologies to more
    complex multi-body floating-base robots, such as humanoids and quadrupeds, as well as
    more in-depth analysis of the convergence behavior of the algorithms proposed in the
    current work.

\paragraph{Acknowledgements}
This material is based upon work supported by the National Science Foundation Graduate
Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science Foundation

\printbibliography

\end{document}